---
title: "SQL Query - missing data
output: html_notebook
---

# Get essentials 

```{r}
library(sqldf)

RR_data <-readxl::read_xlsx("data/RR_data_3.xlsx")

```

# Missing data

```{r}

# paper surveys 

sqldf("
SELECT AVG(paper_missing)
FROM RR_data;
") 
      
# web surveys
sqldf("
SELECT AVG(dig_missing)
FROM RR_data;
")
      
```

Paper surveys have an avergae of 21% missing data while web surveys have an avergae of 7% data.


# Missing data per feature - paper surveys 

## Missing data for random assignment

```{r}

# Yes
sqldf("
SELECT AVG(paper_missing)
FROM RR_data 
WHERE rand_assign = 'Yes';
")
    
# No  
sqldf("
SELECT AVG(paper_missing)
FROM RR_data 
WHERE rand_assign = 'No';
") 

```

Missing data for paper surveys that used random assignment was 21% while missing data for paper surveys that did not use random assignment was 20%


## missing data for incentives

```{r}

# Yes
sqldf("
SELECT AVG(paper_missing)
FROM RR_data 
WHERE incentives = 'Yes';
") #

# No
sqldf("
SELECT AVG(paper_missing)
FROM RR_data 
WHERE incentives = 'No';
") 
```


Missing data for paper surveys that used incentives was 18%, while missing data for paper surveys that did not use incentives was 68%.

## missing data for reminders


```{r}

# Yes
sqldf("
SELECT AVG(paper_missing)
FROM RR_data 
WHERE  reminder = 'Yes';
") 

# No      
sqldf("
SELECT AVG(paper_missing)
FROM RR_data 
WHERE reminder = 'No';
")

```


Missing data for paper surveys that used reminder was 28%, while missing data for paper surveys that did not use reminder was 7%.



# missing data per feature - digital surveys 

## missing data for random assignment

```{r}

# Yes
sqldf("
SELECT AVG(dig_missing)
FROM RR_data 
WHERE rand_assign = 'Yes';
") 

# No
sqldf("
SELECT AVG(dig_missing)
FROM RR_data 
WHERE rand_assign = 'No';
") 

```

Missing data for the digital surveys that used random assignment was 9%, while missing data for digital surveys that did not use random assignment was 65%.


## missing data for reminders


```{r}

# Yes
sqldf("
SELECT AVG(dig_missing)
FROM RR_data 
WHERE reminder = 'Yes';
") 

# No
sqldf("
SELECT AVG(dig_missing)
FROM RR_data 
WHERE reminder= 'No';
") 

```

Missing data for digital surveys that used reminders was 6%, while  missing data for digital surveys that did not use reminder was 9%.

## missing data for incentives

```{r}

# Yes
sqldf("
SELECT AVG(dig_missing)
FROM RR_data 
WHERE incentives = 'Yes';
")

# No
sqldf("
SELECT AVG(dig_missing)
FROM RR_data 
WHERE incentives= 'No';
") 

```

Missing data for digital surveys that used reminders was 8%, while missing data for digital surveys that did not use reminder was 2%.






