---
title: "SQL Query - survey response"
output: html_notebook
---


# Get essentials 

```{r}
library(sqldf)

RR_data <-readxl::read_xlsx("data/RR_data_3.xlsx")

```

# Response rate 

```{r}

# paper surveys 

sqldf("
SELECT AVG(Paper_surv)
FROM RR_data;
") 
      
# web surveys
sqldf("
SELECT AVG(dig_survey)
FROM RR_data;
")
      
```

Paper surveys have an avergae of 67% while web surveys have an avergae of 55% response rate.


# Response rate per feature - paper surveys 

## response rate for random assignment

```{r}

# Yes
sqldf("
SELECT AVG(Paper_surv)
FROM RR_data 
WHERE rand_assign = 'Yes';
")
    
# No  
sqldf("
SELECT AVG(Paper_surv)
FROM RR_data 
WHERE rand_assign = 'No';
") 

```

Response rate for paper surveys that used random assignment was 65% while response rate for paper surveys that did not use random assignment was 72%


## response rate for incentives

```{r}

# Yes
sqldf("
SELECT AVG(Paper_surv)
FROM RR_data 
WHERE incentives = 'Yes';
") #

# No
sqldf("
SELECT AVG(Paper_surv)
FROM RR_data 
WHERE incentives = 'No';
") 
```


The response rate for paper surveys that used incentives was 66%, while the response rate for paper surveys that did not use incentives was 68%.

## response rate for reminders


```{r}

# Yes
sqldf("
SELECT AVG(Paper_surv)
FROM RR_data 
WHERE  reminder = 'Yes';
") 

# No      
sqldf("
SELECT AVG(Paper_surv)
FROM RR_data 
WHERE reminder = 'No';
")

```


The response rate for paper surveys that used reminder was 67%, while the response rate for paper surveys that did not use reminder was 68%.



# Response rate per feature - digital surveys 

## response rate for random assignment

```{r}

# Yes
sqldf("
SELECT AVG(dig_survey)
FROM RR_data 
WHERE rand_assign = 'Yes';
") 

# No
sqldf("
SELECT AVG(dig_survey)
FROM RR_data 
WHERE rand_assign = 'No';
") 

```

The response rate for the digital surveys that used random assignment was 50%, while the response rate for digital surveys that did not use random assignment was 65%.


## response rate for reminders


```{r}

# Yes
sqldf("
SELECT AVG(dig_survey)
FROM RR_data 
WHERE reminder = 'Yes';
") 

# No
sqldf("
SELECT AVG(dig_survey)
FROM RR_data 
WHERE reminder= 'No';
") 

```

The response rate for digital surveys that used reminders was 53%, while  response rate for digital surveys that did not use reminder was 61%.

## response rate for incentives

```{r}

# Yes
sqldf("
SELECT AVG(dig_survey)
FROM RR_data 
WHERE incentives = 'Yes';
")

# No
sqldf("
SELECT AVG(dig_survey)
FROM RR_data 
WHERE incentives= 'No';
") 

```

The response rate for digital surveys that used reminders was 60%, while the response rate for digital surveys that did not use reminder was 52%.






